---
title: "African Elephant Populations"
author: "Camryn Allen"
date: March 27, 2024
output: 
  beamer_presentation:
    theme: "Malmoe" 
    colortheme: "beaver"
fontsize: 11pt
---

```{r, eval=FALSE, echo=FALSE} 
knitr::opts_chunk$set( #what does eval & echo = FALSE mean? 
  echo = FALSE,
  message = FALSE, #meaning?
  warning = FALSE,#meaning?
  prompt = FALSE, #meaning? 
  tidy = FALSE, #meaning?
  comment = NA, #meaning? 
  fig.align = 'center'
  )

library(tidyverse)
library(knitr)
library(rmdformats)
library(readr)
library(ggplot2)
library(patchwork)
```

# The Data Details

Collected by the African Elephant Specialist Group (AfESG) of the Food and Agriculture Organization of the United Nations, the data utilized for this presentation visualizes and compares the elephant populations of different African countries at intervals between 1979 and 2015.


Overview

- Africa's Elephant Population (1979 - 2015)
- Top 10 Largest African Elephant Populations (2015)


# A Focus on Africa's Elephant Population Overtime
```{r, echo=FALSE, eval=TRUE, warning=FALSE,message=FALSE, prompt=FALSE, comment=NA} 
library(dplyr)
library(readr)
african_elephants <- read_csv('african-elephants.csv')

df2 = african_elephants

AfricaElephants = 
  african_elephants %>%
  filter(Entity %in% c('Africa'))

AfricaElephants <- subset(AfricaElephants, Year > 1913)

AfrMostElephants = 
  african_elephants %>%
  filter(Entity %in% c('Botswana', 'Tanzania', 'Zimbabwe'))

AfrMostElephants <- AfrMostElephants %>%
  filter(Year != '2007')

#Plot of African Elephants
library(ggplot2)
AfricaPlot = 
  AfricaElephants |>
  ggplot(aes(x = Year, y = `African Elephant population (AfESG, 2019)`)) + 
  geom_line(
    color = '#18A0F4'
  ) + 
  ggtitle('African Elephant Population (1979 - 2015)') + 
  labs(
    x = 'Year',
    y = "Elephant Population"
  ) +
  scale_x_continuous(
    breaks = c(1970,1980,1990,2000,2010)
  ) + 
  scale_y_continuous(
    labels = c('250,000', '500,000', '750,000', '1.0M', '12.5M')
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )

AfricaPlot
```

# Top 10 Largest African Elephant Populations (2015)

```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
library(readr)
library(dplyr)

d3 = african_elephants

d3 <- d3 %>%
  filter(
    Year == '2015',
    `African Elephant population (AfESG, 2019)` != '415428', 
    `African Elephant population (AfESG, 2019)` != '24119', 
    `African Elephant population (AfESG, 2019)` != '86373', 
    `African Elephant population (AfESG, 2019)` != '293447', 
    `African Elephant population (AfESG, 2019)` != '11489'
    )

PleaseLord = d3
PleaseLord$Entity <- reorder(PleaseLord$Entity, +PleaseLord$`African Elephant population (AfESG, 2019)`)

PrettyPlease = 
  PleaseLord %>%
  filter(Entity %in% c('Botswana', 'Zimbabwe', 'Tanzania', "Kenya", 'Namibia', 'Zambia', 'South Africa', 'Mozambique', 'South Sudan', 'Gabon'))

Africa2015Plot = 
 PrettyPlease |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity)) + 
  geom_bar(stat = 'identity', width = 0.8, fill = '#18A0F4') + 
  ggtitle('African Elephant Populations in 10 Countries (2015)') + 
  labs(
    x = 'Elephant Population',
    y = "Country"
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K'),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )

Africa2015Plot
```

# Highest Elephant-Populated Regions: Eastern and Southern Africa
```{r, eval=TRUE, echo=FALSE}
DensEl = 
  african_elephants %>%
  filter(Entity %in% c('Eastern Africa', 'Southern Africa'))

DensEl$Density <- DensEl$`African Elephant population (AfESG, 2019)` / 1000

DPlot = 
  DensEl |>
  ggplot(aes(y = Density, x = Year)) +
  geom_area(alpha = 0.70, color = 'black', fill = '#18A0F4') + 
  labs(
    title = 'Elephants in Eastern and Southern Africa (1995 - 2015)',
    y = 'Elephant Population (thousands)',
    x = 'Year'
  ) + 
   scale_x_continuous(
    expand = expansion(mult = c(0, 0.08)),
    limits = c(1995, 2015)
    ) + 
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)
  )) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )
  
  
  DPlot
```

# A Comparison of African Elephant Populations 
```{r, echo=FALSE, eval=TRUE, fig.width=10, fig.height=6, warning=FALSE}
library(patchwork)
library(ggplot2)

d3 = african_elephants

d3 <- d3 %>%
  filter(
    Year == '2015',
    `African Elephant population (AfESG, 2019)` != '415428', 
    `African Elephant population (AfESG, 2019)` != '24119', 
    `African Elephant population (AfESG, 2019)` != '86373', 
    `African Elephant population (AfESG, 2019)` != '293447', 
    `African Elephant population (AfESG, 2019)` != '11489'
    )
  
PleaseLord = d3

PleaseLord = d3
PleaseLord$Entity <- reorder(PleaseLord$Entity, +PleaseLord$`African Elephant population (AfESG, 2019)`)

PrettyPlease = 
  PleaseLord %>%
  filter(Entity %in% c('Botswana', 'Zimbabwe', 'Tanzania', "Kenya", 'Namibia', 'Zambia', 'South Africa', 'Mozambique', 'South Sudan', 'Gabon'))

Africa2015Plot = 
 PrettyPlease |>
  ggplot(aes(x = `African Elephant population (AfESG, 2019)`, y = Entity)) + 
  geom_bar(stat = 'identity', width = 0.8, fill = '#18A0F4') + 
  labs(
    title = 'Top 10 African Elephant Populations',
    x = 'Elephant Population',
    y = "Country",
    fill = 'Approximate Population'
  ) + 
  scale_x_continuous(
    breaks = c(5000,25000,50000,75000,100000,125000),
    labels = c('5K', '25K', '50K', '75K', '100K', '125K'),
    expand = expansion(mult = c(0, 0.05))
  ) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold'),
    legend.position = "none"
  )

DPlot = 
  DensEl |>
  ggplot(aes(y = Density, x = Year)) +
  geom_area(alpha = 0.70, color = 'black', fill = '#18A0F4') + 
  labs(
    title = 'Elephants in Eastern and Southern Africa',
    y = 'Elephant Population (thousands)',
    x = 'Year'
  ) + 
   scale_x_continuous(
    expand = expansion(mult = c(0, 0.08)),
    limits = c(1995, 2015)
    ) + 
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1)
  )) +
  theme_bw() + 
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = 'bold'),
    axis.title.x = element_text(face = 'bold'),
    axis.title.y = element_text(face = 'bold')
  )

Africa2015Plot + DPlot + 
  plot_annotation(
    theme = theme(plot.title = element_text(size = 12, face = 'bold', hjust = 0.5)))

```



# Extra Credit

Special Easter Egg from PS5 key.Rmd: "I'm a very special easter egg in that I only appear in the markdown file. Did you find me? Email Hart for your 3/27 extra credit!"
